Fusing Filters with Integer Linear Programming

### 2014-05-12 2014betterfusionforfilters/section/04-ILP.tex
start fixing ILP different types

\paragraph{Fusion between different types}
\ben{We haven't said what a ``type'' is in the context of clustering. Do we mean ``rate''? If so, can we use $k$ for the variables instead of $\tau$?}

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge trans_i=\emptyset \wedge trans_j=\emptyset$)                 \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge trans_i=\{i'\}$) \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge trans_j=\{j'\}$) \\
           \> \ldots
\end{tabbing}
The main difference to existing integer linear programming solutions is that, while nodes with different loop sizes cannot generally be fused together, we do support some fusion of filtered sizes. If two nodes have different sizes or rates, but are both sub-rates of another rate, they \emph{can} be fused together if they are both fused with their generators. \ben{What is a subrate? How do we determine the subrates of each rate?} \ben{Highlight why we want to allow this sort of fusion, and why it is specific to data flow fusion.} \ben{Explain how we are using the idea of transducers}.

>>>>

\paragraph{Fusion between different iteration sizes}

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\tau(i) \not= \tau(j) \wedge parents(i,j) = \emptyset$)  \\
           \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> $x_{i'j'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau(i) \not= \tau(j) \wedge parents(i,j) = \{i',j'\}$) \\
           \> \ldots
\end{tabbing}
If two nodes have different iteration sizes, they may not necessarily be fused.
However, if they have parent transducers of the same size, they may be fused if both their parents are fused together, and each is fused with its respective parent.

The main difference to existing integer linear programming solutions is that, while nodes with different loop sizes cannot generally be fused together, we do support some fusion of filtered sizes. If two nodes have different sizes or rates, but are both sub-rates of another rate, they \emph{can} be fused together if they are both fused with their generators. \ben{What is a subrate? How do we determine the subrates of each rate?} \ben{Highlight why we want to allow this sort of fusion, and why it is specific to data flow fusion.} \ben{Explain how we are using the idea of transducers}.

As a degenerate example, consider fusing an operation on filtered data with its generating filter:
\begin{code}
gts  = filter (>0) us
sum2 = fold (+) 0  gts
\end{code}
Here, $parents(gts,sum2) = \{gts, gts\}$.
This generates spurious, but still valid, constraints that $gts$ must be fused with $gts$ and $gts$ must be fused with $sum2$, in order for $gts$ and $sum2$ to be fused together.
While these constraints are unnecessary in this case, they are harmless.
